### 2015-01-20 cover.tex
cover letter

\documentclass{letter}
\usepackage{fullpage}
\signature{Nancy Soontiens}
\address{Department of Earth, Ocean, and Atmospheric Sciences\\ University of British Columbia \\ Vancouver, BC
\\ nsoontie@eos.ubc.ca}
\begin{document}

\begin{letter}{}

\opening{Dear Editor,}

On my behalf of my co-authors I am submitting an article titled ``Storm surges in the Strait of Georgia simulated with a regional model" for consideration for publication in Atmosphere-Ocean. This article presents research undertaken by scientists collaborating across Canada to produce a new regional ocean model of the Salish Sea.  We believe that this topic is interesting and relevant to the readership of Atmosphere-Ocean, particularly those readers engaged in regional ocean modelling in Canada.

I look forward to hearing from you at your earliest convenience.

\closing{Sincerely,}
\end{letter}
\end{document}